---
title: "Assignment10-.2"
author: "Joshua Burden"
date: "`r Sys.Date()`"
output:
  pdf_document: default
  html_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggm)
library(ggplot2)
library(readr)
library(pastecs)
library(plyr)
library(dplyr)
library(magrittr)
library(purrr)
library(ggplot2)
library(knitr)
```
 
```{r include=TRUE, , echo=FALSE, message = FALSE, warning = FALSE}
setwd("/Users/joshua/Documents/PERSONAL_GITHUB_REPOS/dsc520")
## Read housing dataset
TS_df <- read.csv("data/thoraricSurgery.csv")
```

## Fit a Logistic Regression Model to Thoracic Surgery Binary Dataset

- Fit a binary logistic regression model to the data set that predicts whether or not the patient survived for one year (the Risk1Y variable) after the surgery. 
- Use the glm() function to perform the logistic regression. 
- See Generalized Linear Models for an example.
- Include a summary using the summary() function in your results.

```{r} 
TS_df$DGN <- as.factor(TS_df$DGN)
TS_df$PRE14 <- as.factor(TS_df$PRE14)
thoracicSurgery <- glm(Risk1Yr ~  DGN + PRE4 + PRE5 + PRE6 + PRE7 + PRE8 + PRE9 + PRE10 + PRE11 + PRE14 + PRE17 + PRE19 + PRE25 + PRE30 + PRE32 + AGE, data = TS_df, family = binomial())
summary(thoracicSurgery)
```

### According to the summary, which variables had the greatest effect on the survival rate?

> PRE9, PRE14OC14 has the greatest effect on the survival rate.

- To compute the accuracy of your model, use the dataset to predict the outcome variable. 
- The percent of correct predictions is the accuracy of your model. 
- What is the accuracy of your model?

```{r}
predict_data <- predict(thoracicSurgery,TS_df,type="response") > .5
totalCorrect <- sum( (TS_df$Risk1Yr == "T") == (predict_data) )
totalPercent <- round(sum( (TS_df$Risk1Yr == "T") == (predict_data) ) / nrow(TS_df),2)*100
totalRows <- nrow(TS_df)
totalWrong <- totalRows - totalCorrect
print(paste0("The model predicted ",totalCorrect," successful outcomes. Out of ",totalRows," elements, ",totalWrong, " elements were found incorrect."))
print(paste0("Accruacy score: ",round(totalPercent,digits = 2),"%"))
```


## Fit a Logistic Regression Model
- Fit a logistic regression model to the binary-classifier-data.csv dataset
- The dataset (found in binary-classifier-data.csv) contains three variables; label, x, and y.
- The label variable is either 0 or 1 and is the output we want to predict using the x and y variables.
```{r}
setwd("/Users/joshua/Documents/PERSONAL_GITHUB_REPOS/dsc520")
bc_df <- read_csv("data/binary-classifier-data.csv")
ggplot(bc_df, mapping = aes(x = x, y = y)) + geom_point(mapping = aes(colour = y)) 
```

### What is the accuracy of the logistic regression classifier?
```{r logistic accuracy, include=TRUE}
model1_xy <- glm(label ~ x + y, data = bc_df, family = "binomial")
model1_x  <- glm(label ~ x, data = bc_df, family = "binomial")

summary(model1_xy)
summary(model1_x)
```



Keep this assignment handy, as you will be comparing your results from this week to next week.